---
title: "p8105_hw2_yz3297"
author: "Yue Zhao"
date: "10/03/2018"
output: github_document
---
##Problem 1
```{r setup, include=FALSE}
library(tidyverse)

```

```{r, results='hide'}
  #read, clean, retain variables and change "entry" from character to logical
  subway_data = 
    read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
    janitor::clean_names()  %>%
    select(line, station_name, station_latitude, station_longitude, route1:route11 , entry, vending, entrance_type, ada) %>%
    mutate(entry= if_else(entry == "YES", TRUE, FALSE))
    
``` 

This code chunk was to read, clean, retain variables and change "entry" from character to logical.

The dataset has 19 variables which are Line, Station_name, Station_latitude, Station_longitude, route1-route11 , Entry, Vending, Entrance_type, ADA compliance. 
It is 1868*19 in dimension. This dataset is obviously not tidy because route has multiple variables. We need to transform the dataset from wide to long form.


```{r}
    count(distinct(subway_data,station_name, line)) 
```

There are `r count(distinct(subway_data,station_name, line))` distinct stations. (identified by name and line)


```{r}
    count(distinct(subway_data,station_name, line, ada), ada== TRUE)
```  


There are 84 distinct stations which are ADA compliant.

```{r}
    count(subway_data, vending == "NO", entry == TRUE)
    count(subway_data, vending == "NO")

```   

The proportion of station entrances without vending allow entrance is 69/183 `r 69/183`.


```{r}
    subway_data = gather(subway_data, key = route_number , value = route_name, route1:route11)
```

The code chunk above is to change the dataset from wide form to long form. 


```{r}
    count(distinct(subway_data, station_name, line, route_name), route_name == "A")
    count(distinct(subway_data, station_name, line, route_name, ada), route_name == "A" , ada == TRUE)
    
```

There are 60 distinct stations serve the A train. Of the stations that serve the A train, 17 of them are ADA compliant. 


##Problem 2
```{r}
library(readxl)

  trashwheel_data = 
      read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "Mr. Trash Wheel" , range = "A2:N337") %>%
      janitor::clean_names()  %>%
      rename(weight = weight_tons, volume = volume_cubic_yards)   %>%
      #read, clean names and rename weight and volume
      
      na.omit()  %>%
      #omit the redundant observations
    
      mutate(sports_balls = as.integer(sports_balls))
      #convert sports balls to an integer variable
  
  trashwheel_data
      
```

This code chunk is to read, clean names, rename weight and volume, omit the redundant observations and convert sports balls to an integer variable for Mr. Trash Wheel dataset. 


```{r}     
    precip2017_data=
      read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation" , range = "A2:B15") %>%
      na.omit() %>%
      mutate(year = 2017)
      #read, clean data and add year= 2017
      
    precip2016_data=  
      read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2016 Precipitation" , range = "A2:B15") %>%
      na.omit() %>% 
      mutate(year = 2016)
      #read, clean data and add year= 2016
      
    precip_data=
      bind_rows(precip2017_data, precip2016_data) %>%
      janitor::clean_names()  %>%
      mutate(month = month.name[month])
      #append the datasets, clean the names and convert month to a character variable
  
    precip_data
   
```

This code chunk is to read the precipitation datasets and combine the datasets. 


The description about the data (r inline code used): The number of observations in Mr. Trashwheel Dataset is `r count(trashwheel_data)`. The number of observations in the Precipitation Dataset is `r count(precip_data)`. 

Key variables in Mr. Trashwheel Dataset are dumpster, month, weight, volume, sports balls. Key variables in Precipitation Dataset are month, total precipitation, year. 

The total precipitation in 2017 is `r sum(precip2017_data$Total)`. The median number of sports balls in a dumpster in 2016 is `r median(filter(trashwheel_data,year== 2016)$sports_balls)`.


##Problem 3

```{r}  
devtools::install_github("p8105/p8105.datasets")

library(p8105.datasets)

data(brfss_smart2010)
head(brfss_smart2010)

brfssnew_data=brfss_smart2010 %>%
  janitor::clean_names()  %>%
  filter(topic=="Overall Health")   %>%
  select(year,locationabbr, locationdesc, response, data_value)%>%
  spread(key=response, value= data_value ) %>%
  #clean the names, filter the overall health topic, drop other variables and turning data from long to wide
  
  janitor::clean_names()  %>%
  mutate(exc_or_verygood= excellent + very_good ) %>%
  #clean the names again after turning long to wide and create a new variable

  na.omit()
  #omit the na

brfssnew_data
```


Answers to the questions (used inline code for the numbers if no code chunk is showing)

There are `r count(distinct(brfssnew_data, locationdesc))` unique locations in the dataset (if counting the NAs, it is 404).  There are `r count(distinct(brfssnew_data, locationabbr))` location abbreviations for states (or district), so we know all 50 states and Washington DC were represented.  

```{r}
count(brfssnew_data, locationabbr)
```

The state which has the most observation is New Jersey. It appeared `r max(count(brfssnew_data, locationabbr)$n)` times. (The two NAs are taken out so it could also be 146)

In 2002, the median of the Excellent response value is `r median(filter(brfssnew_data,year==2002)$excellent)`.

```{r}
ggplot(filter(brfssnew_data,year==2002), aes(x = excellent)) + 
  geom_histogram()
  #histogram
```

Histogram of Excellent response values in 2002

```{r}
  county_data1= filter(brfssnew_data, locationdesc=="NY - New York County") %>%
            select(year, excellent) %>%
            rename(excellent_NY = excellent)
  #subseting the New York County data

  county_data2 = filter(brfssnew_data, locationdesc=="NY - Queens County")%>%
            select(year, excellent) %>%
            rename(excellent_QC = excellent)
  #subseting the Queens County data
  
  county_data =left_join(county_data1,county_data2, by = "year")
  #combine the data
       
  ggplot(county_data, aes(x = excellent_NY, y = excellent_QC)) + geom_point()
  #scatter plot
       
```

Scatter Plot of Execellent response values in New York County and Queens County from 2002 to 2010. (There are 9 points)











