---
title: "p8105_hw2_yz3297"
author: "Yue Zhao"
date: "10/03/2018"
output: github_document
---
##Problem 1
```{r setup, include=FALSE}
library(tidyverse)
```

```{r, results='hide'}

  subway_data = 
    read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
    janitor::clean_names()  %>%
    select(line, station_name, station_latitude, station_longitude, route1:route11 , entry, vending, entrance_type, ada) %>%
    mutate(entry= if_else(entry == "YES", TRUE, FALSE))
    
``` 

This code chunk was to read, clean, retain variables and change "entry" from character to logical.

The dataset has 19 variables which are Line, Station_name, Station_latitude, Station_longitude, route1-route11 , Entry, Vending, Entrance_type, ADA compliance. 
It is 1868*19 in dimension. This dataset is obviously not tidy because route has multiple variables. We need to transform the dataset from wide to long form.



```{r}
    count(distinct(subway_data,station_name, line)) 
```

There are 465 distinct stations. (identified by name and line)




```{r}
    count(distinct(subway_data,station_name, line, ada), ada== TRUE)
```  

```{r}
    count(subway_data, vending == "NO", entry == TRUE)
    count(subway_data, vending == "NO")
```     

```{r}
    subway_data = gather(subway_data, key = route_number , value = route_name, route1:route11)
```

```{r}
    count(distinct(subway_data, station_name, line, route_name), route_name == "A")
    count(distinct(subway_data, station_name, line, route_name, ada), route_name == "A" , ada== TRUE)
    
```

